\documentclass[11pt]{article}

\usepackage[margin=1in]{geometry}
\usepackage{enumitem}
\usepackage{hyperref}
\usepackage{setspace}

\setlist[itemize]{leftmargin=1.5em}
\setstretch{1.1}

\title{To--Do List for Applying to Graduate School in Physics\\
Fall 2026 Admission}
\author{}
\date{}

\begin{document}

\maketitle

\section*{Overview}
This document provides a structured timeline and checklist for applying
to graduate programs (PhD or MS) in physics for Fall 2026 admission.
The tasks are organized chronologically to support early preparation,
strong applications, and informed decision--making.

\section*{Now -- Spring 2025: Laying the Groundwork}
\textbf{Goal: Build a strong academic profile and clarify direction.}

\begin{itemize}
  \item Decide which degree(s) to pursue (PhD, MS, or both).
  \item Identify intended subfield(s):
    \begin{itemize}
      \item Condensed matter, AMO, high energy physics, astrophysics,
            biophysics, computational physics, etc.
    \end{itemize}
  \item Meet with faculty mentors to discuss graduate school plans and competitiveness.
  \item Begin or deepen involvement in a research project.
  \item Maintain an up--to--date CV including:
    \begin{itemize}
      \item Research experience
      \item Presentations or posters
      \item Programming skills
      \item Teaching or mentoring experience
    \end{itemize}
  \item Create a centralized document or notebook to track:
    \begin{itemize}
      \item Potential programs
      \item Deadlines
      \item Requirements
      \item Faculty of interest
    \end{itemize}
\end{itemize}

\section*{Summer 2025: Exploration and Preparation}
\textbf{Goal: Narrow focus and prepare application components.}

\begin{itemize}
  \item Compile a long list of graduate programs (15--25).
  \item Identify potential research advisors at each institution.
  \item Read recent publications by faculty of interest.
  \item Decide whether to take the GRE or Physics GRE (if applicable).
  \item Begin test preparation if required.
  \item Engage in full--time research if possible (REU, thesis, or summer project).
\end{itemize}

\section*{Fall 2025: Application Materials Phase I}
\textbf{Goal: Secure recommenders and draft core documents.}

\begin{itemize}
  \item Narrow the list of schools to 8--15 programs.
  \item Identify three letter writers:
    \begin{itemize}
      \item Primary research advisor
      \item Additional faculty familiar with academic work
      \item Teaching or computational mentor (if relevant)
    \end{itemize}
  \item Request letters of recommendation early (September).
  \item Provide letter writers with:
    \begin{itemize}
      \item CV
      \item Unofficial transcripts
      \item Draft statements
      \item List of deadlines
    \end{itemize}
  \item Draft a general Statement of Purpose.
  \item Draft a Personal Statement (if required).
  \item Obtain and review unofficial transcripts for accuracy.
\end{itemize}

\section*{Winter 2025--2026: Application Materials Phase II}
\textbf{Goal: Polish and customize applications.}

\begin{itemize}
  \item Customize Statements of Purpose for each program:
    \begin{itemize}
      \item Faculty fit
      \item Departmental strengths
      \item Research alignment
    \end{itemize}
  \item Finalize CV formatting and content.
  \item Take GRE or Physics GRE if applicable.
  \item Create a master application spreadsheet tracking:
    \begin{itemize}
      \item Program name
      \item Application deadline
      \item Letter submission status
      \item Application fees
      \item Submission status
    \end{itemize}
\end{itemize}

\section*{Winter 2025: Application Submission (October--December)}
\textbf{Goal: Submit complete, high--quality applications on time.}

\begin{itemize}
  \item Submit applications at least 1--2 weeks before deadlines.
  \item Confirm letters of recommendation have been submitted.
  \item Apply for fee waivers where eligible.
  \item Verify all required materials:
    \begin{itemize}
      \item Statements
      \item Transcripts
      \item Test scores
      \item Supplementary documents
    \end{itemize}
  \item Save PDF copies of all submitted applications.
\end{itemize}

\section*{Spring 2026: Interviews and Decisions}
\textbf{Goal: Evaluate offers and select a program.}

\begin{itemize}
  \item Prepare for interviews by reviewing:
    \begin{itemize}
      \item Research projects
      \item Faculty interests
      \item Departmental culture
    \end{itemize}
  \item Practice explaining research clearly and concisely.
  \item Attend visit days (virtual or in--person).
  \item Ask questions about:
    \begin{itemize}
      \item Advisor availability
      \item Funding guarantees
      \item Time to degree
      \item Career placement
    \end{itemize}
  \item Compare offers holistically.
  \item Make a final decision by April 15.
\end{itemize}

\section*{Commonly Overlooked Points}
\begin{itemize}
  \item Begin writing statements earlier than expected.
  \item Letters of recommendation often outweigh institutional prestige.
  \item Research fit is more important than rankings.
  \item Keep recommenders informed of outcomes.
  \item Archive finalized Statements of Purpose for future reference.
\end{itemize}

\end{document}
